---
title: "Ryan Tillis - Machine Learning - Data Science - Quiz 1 - Coursera"
author: <a href="http://www.ryantillis.com"> Ryan Tillis </a>
date: "June 3, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Machine Learning Quiz 1

This is Quiz 1 from the Machine Learning course within the Data Science Specialization. This publication is intended as a learning resource, all answers are documented and explained.

###Questions
<hr>
<font size="+2">1.  </font> Which of the following are components in building a machine learning algorithm?

<hr>

* <font size="+1">**Creating features**

</font>

<hr>

#####Explanation:

Self-Explanatory

```{r Question 1}

```

<hr>

<font size="+2">2.  </font> Suppose we build a prediction algorithm on a data set and it is 100% accurate on that data set. Why might the algorithm not work well if we collect a new data set?

<hr>

* <font size="+1">**Our algorithm may be overfitting the training data, predicting both the signal and the noise.**</font>

<hr>
```{r Question 2}

```

#####Explanation:

If your algorithm is too specific to the peculiarities of an individual set, it may lead to false classification and bad predictions.

<hr>

<font size="+2">3.  </font> What are typical sizes for the training and test sets?

<hr>

* <font size="+1"> **60% in the training set, 40% in the testing set.**

</font>

<hr>
```{r Question 3}

```

#####Explanation:

Self-explanatory

<hr>

<font size="+2">4.  </font> What are some common error rates for predicting binary variables (i.e. variables with two possible values like yes/no, disease/normal, clicked/didn't click)? Check the correct answer(s).

<hr>

* <font size="+1">**Predictive value of a positive**
</font>

<hr>
```{r Question 4}

```

#####Explanation:

Self-explanatory

<hr>

<font size="+2">5.  </font> Suppose that we have created a machine learning algorithm that predicts whether a link will be clicked with 99% sensitivity and 99% specificity. The rate the link is clicked is 1/1000 of visits to a website. If we predict the link will be clicked on a specific visit, what is the probability it will actually be clicked?

<hr>

* <font size="+1"> **9%** </font>

<hr>

#####Explanation:


```{r Question 5}
n <- 1000000
prevalence = 1/1000
sens <- .99
spec <- .99
clicked <- prevalence * n
TP <- sens * clicked
Notclicked <- n - clicked
TN <- spec * Notclicked
FP <- Notclicked * sens
FP<- Notclicked-TN
PPV <- TP/ (TP+FP)

PPV
```

<hr>

Check out my website at: <http://www.ryantillis.com/>

<hr>
